---
title: "IPZC3H4 Integrated analysis"
author: "Fabio Bedin | MS-Unit"
output: html_document
---


```{r librerie, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(fig.align = "center", warning=FALSE, message=FALSE, fig.height=8, fig.width=12)
library("MBQN")
library("DEP")
library("SummarizedExperiment")
library("preprocessCore")
library("enrichR")
library("DT")
library("patchwork")
library("here")
library("datapasta")
library("ggstatsplot")
library("UpSetR")
library("ggrepel")
library("visdat")
library("naniar")
library("gprofiler2")
library("openxlsx")
library("org.Hs.eg.db")
library("rrvgo")

library("tidyverse")
```

```{r Custom Functions}
source(here::here("code/custom_functions.R"))
```

```{r excel_define_wb}
header_style <- createStyle(
  fontSize = 12,
  fontColour = "#0f0f0f",
  fgFill = "#faf2ca",
  halign = "center",
  border = "TopBottomLeftRight")

body_style <- createStyle(
  halign = "center",
  border = "TopBottomLeftRight")

excel <- createWorkbook()
```

```{r load_PG}
data <- read.csv(here::here("data/proteinGroups_all.txt"), header = TRUE,stringsAsFactors = FALSE, sep = "\t")

data <- data[data$Reverse != "+" & data$Potential.contaminant != "+" & data$Only.identified.by.site != "+" & data$Peptides >= 2,]

data_unique <- make_unique(data, "Gene.names", "Protein.IDs", delim = ";")
```

```{r expdes}
cond_1 <- "IPZC3H4"

cond_2 <- "IPEV"

conditions<-c(cond_1,cond_2)

test<- paste(cond_1,cond_2,sep="_vs_")

expdesign <- read.table(here::here("data/expdesign_all.tsv"), header = T, stringsAsFactors = F)

expdesign <- subset(expdesign, condition %in% conditions)

columns<-match(paste("LFQ.intensity.",expdesign$label,sep=""),colnames(data_unique))

data_se <- make_se(data_unique, columns, expdesign)

data_filt <- filter_missval(data_se, thr = 1)
```

## Quality control plots {.tabset .tabset-fade}

### Proteins numbers

```{r proteins_1}
plot_numbers_lables(data_filt)
```

### normalization

```{r normalization_1}
plot_normalization(data_filt)
```

### upset plot {.tabset .tabset-fade}

#### `r cond_1`

```{r upset_1.1}
assay(data_filt) %>% 
  as.data.frame() %>% 
  select(starts_with(cond_1)) %>% 
  rownames_to_column() %>% 
  pivot_longer(!rowname, names_to = "samples", values_to = "intensity") %>% 
  mutate(intensity = if_else(is.na(intensity), 0, 1)) %>% 
  pivot_wider(names_from = samples, values_from = intensity) %>% 
  as.data.frame() %>% 
  upset(nsets = 5,
        order.by = "freq",
        keep.order = T,
        text.scale = 2.5,
        point.size = 4,
        line.size = 0.5, 
        sets.bar.color = "coral2",
        main.bar.color  = "gray44")
```

#### `r cond_2`

```{r upset_1.2}
assay(data_filt) %>% 
  as.data.frame() %>% 
  select(starts_with(cond_2)) %>% 
  rownames_to_column() %>% 
  pivot_longer(!rowname, names_to = "samples", values_to = "intensity") %>% 
  mutate(intensity = if_else(is.na(intensity), 0, 1)) %>% 
  pivot_wider(names_from = samples, values_from = intensity) %>% 
  as.data.frame() %>% 
  upset(nsets = 5,
        order.by = "freq",
        keep.order = T,
        text.scale = 2.5,
        point.size = 4,
        line.size = 0.5, 
        sets.bar.color = "turquoise3",
        main.bar.color  = "gray44")
```

#### Both

```{r upset_1.3}
assay(data_filt) %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  pivot_longer(!rowname, names_to = "samples", values_to = "intensity") %>% 
  mutate(intensity = if_else(is.na(intensity), 0, 1)) %>% 
  pivot_wider(names_from = samples, values_from = intensity) %>% 
  as.data.frame() %>% 
  upset(nsets = 610,
        order.by = "freq",
        keep.order = T,
        text.scale = 2.5,
        point.size = 4,
        line.size = 0.5, 
        sets.bar.color = rep(c("turquoise3", "coral2"), each = 5),
        main.bar.color  = "gray44")
```


## DEP analysis {.tabset .tabset-fade}

```{r diff_1}
data_diff <- test_diff_BH(data_filt, type = "manual", test = test)

dep <- add_rejections(data_diff, alpha = 0.05, lfc = 1)

results<-get_results(dep)
```

```{r excel_res_table_1}
a <- get_df_wide(dep) %>% 
  select(c(name, Protein.IDs, Protein.names, starts_with(conditions), -ends_with(c("CI.R", "CI.L")), starts_with("Peptides."))) %>% 
  dplyr::rename_with(., ~ gsub(pattern = paste0(test, "_"), replacement = "", .), starts_with(test)) %>% 
  dplyr::rename_with(., ~ gsub(pattern = "^", replacement = "LFQ_intensity_", .), starts_with(conditions)) %>%
  dplyr::rename(FC = diff) %>% 
  mutate(significant = if_else(significant, "+", "")) %>%
  arrange(desc(significant), p.val) %>% 
  mutate(across(c(p.adj, p.val), format, scientific = T, digits = 2)) %>% 
  mutate(across(starts_with(c("LFQ", "FC")), .fns = ~ round(.x, digits = 2))) %>% 
  relocate(significant) %>%
  relocate(starts_with("LFQ"), .after = p.val)

addWorksheet(excel, sheetName = test, gridLines = F)

writeDataTable(excel, sheet = test, x = a, keepNA = T, na.string = "NaN")

n_row <- a %>% nrow() + 1

n_col <- a %>% ncol()

setColWidths(excel, sheet = test, cols = 1:n_col, widths = 21)

addStyle(excel, sheet = test, style = header_style, rows = 1, cols = 1:n_col, gridExpand = T)

addStyle(excel, sheet = test, style = body_style, rows = 2:n_row, cols = 1:n_col, gridExpand = T)

```

There are **`r results %>% filter(significant) %>% nrow()`** significant proteins.

### PCA

```{r PCA_1}
plot_pca(dep, x = 1, y = 2, n = 200, point_size = 4, indicate = c("condition", "replicate"))
```

### Volcano 

```{r volcano_1}
plot_volcano(dep, contrast=test, add_names=T,label_size=5,adjusted = F)
```

### Result table

```{r res_teble}
results %>% 
  select(starts_with("name") | starts_with(test)) %>% 
  filter(across(ends_with("significant"))) %>% 
  arrange(across(ends_with("p.adj"))) %>%
  mutate(across(2:3, format, scientific = T, digits = 2)) %>%
  dplyr::rename_with(~ tolower(gsub(paste0(test,"_"), "", .x)), .cols = starts_with(test)) %>% 
  mutate(significant = str_to_title(significant)) %>%
  dplyr::rename(FC = ratio) %>% 
  DT::datatable(options = list(
  columnDefs = list(list(className = 'dt-center', targets = 1:5)),
  pageLength = 10)) %>%
  formatStyle('FC', backgroundColor = styleInterval(c(-1, 1), c('lightblue', 'white', 'tomato'))) %>% 
  formatStyle('significant', color = styleEqual(c("True", "False"), c('green', 'red')))
```


### Unique

```{r unique_1}
data_filt_1 <- filter_missval(data_se, thr = 1)

unique_pair(data_filt_1, conditions = conditions)

a <- unique_pair(data_filt_1, conditions = conditions, table = T)

b <- a %>% 
  pivot_longer(!starts_with(c("name", "unique_", "Protein.")), names_to = "samples", values_to = "intensity") %>% 
  mutate(cond = gsub(pattern = "..$", "", samples)) %>% 
  mutate(intensity = case_when(is.na(intensity) ~ 0, TRUE ~ as.numeric(1))) %>%
  group_by(name, cond) %>% 
  mutate(Unique_class_B = case_when(sum(intensity) == 4 ~ "+")) %>% 
  mutate(Unique_class_A = case_when(sum(intensity) == 5 ~ "+")) %>% 
  ungroup() %>% 
  select(-cond) %>% 
  group_by(name) %>% 
  fill(Unique_class_B, .direction = "downup") %>% 
  fill(Unique_class_A, .direction = "downup") %>% 
  ungroup() %>% 
  pivot_wider(names_from = samples, values_from = intensity) %>% 
  select(name, Unique_class_B, Unique_class_A) %>% 
  mutate(across(starts_with("Unique_"), ~ case_when(is.na(.x) ~ "", TRUE ~ as.character(.x))))

a <- a %>% left_join(b)

unique_name <- paste0("Unique_", test)

addWorksheet(excel, sheetName = unique_name, gridLines = F)

writeDataTable(excel, sheet = unique_name, x = a, keepNA = T, na.string = "NaN")

n_row <- a %>% nrow() + 1

n_col <- a %>% ncol()

setColWidths(excel, sheet = unique_name, cols = 1:n_col, widths = 21)

addStyle(excel, sheet = unique_name, style = header_style, rows = 1, cols = 1:n_col, gridExpand = T)

addStyle(excel, sheet = unique_name, style = body_style, rows = 2:n_row, cols = 1:n_col, gridExpand = T)
```

## Gene Ontology {.tabset .tabset-fade}

Then we perfom a gene onology analysis (GO or by branch GO:MF, GO:BP, GO:CC) and a KEGG ontology.

The enrichment results are visualized with an interactive **Manhattan-like-plot**:

### IPZC3H4

Both significant and unique proteins identify are used for GO:

```{r GO_1}
Gname <- results %>% filter(significant & get(paste0(test, "_ratio")) > 0) %>% pull(name)

Uni <- a  %>% filter(unique_IPZC3H4 == "+") %>% pull(name) %>% unique()

GO_name <- c(Gname, Uni)

GO <- gost(query = GO_name, organism = "hsapiens", sources = c("GO:BP", "GO:MF", "GO:CC", "KEGG"), ordered_query = T)

gostplot(GO, capped = TRUE, interactive = TRUE)

write.table(GO_name, here::here("output/Significant.txt"), quote = F, sep = "\t", col.names = F, row.names = F)
```

### Table results

The results can also be visualized with a table: 

```{r GO_table_1}
GO$result %>% 
  as.data.frame() %>% 
  select(starts_with(c("source", "term_name", "p_value", "term_size", "intersection_size"))) %>% 
  arrange(p_value) %>%  
  mutate(p_value = format(p_value, scientific = T, digits = 2)) %>% 
  DT::datatable(options = list(
  columnDefs = list(list(className = 'dt-center', targets = 1:6)),
  pageLength = 10)) %>%
  formatStyle('source', backgroundColor = styleEqual(c("GO:BP", "GO:MF", "GO:CC", "KEGG"), c('orange', 'tomato', 'lightgreen', 'violet')))
```


## {-}

```{r excel}
saveWorkbook(excel, here::here("output/results_GN_MR_IPZC3H4_210721.xlsx"), overwrite = T)
```
